Calibrating seed-based alignment heuristics with Sesame

### 2019-12-10 manuscript.tex
Add section about Sesame to the manuscript

\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{blkarray}
\usepackage[font=small]{caption}
\usepackage{cite}
\usepackage{enumerate}
\usepackage{graphicx}

@@

\title{Calibrating seed-based heuristics to map short DNA reads}

@@

\begin{abstract}
% Statement of the problem.
The increasing throughput of DNA sequencing technologies creates a need
for faster algorithms. The fate of most reads is to be mapped to a
reference sequence, typically a genome. Modern mappers rely on heuristics
to gain speed at a reasonable cost for accuracy. In the seeding approach,
short matches between the reads and the genome are used to narrow the
search to a set of candidate locations. Several seeding variants used in
modern mappers show good empirical performance but they are difficult to
calibrate or to optimize for lack of theoretical results.
% Summary of the results.
Here we develop a theory to estimate the probability that reads are
mapped to a wrong location due to limitations at the seeding step. We
describe the properties of simple exact seeds, skip-seeds and MEM seeds
(Maximal Exact Match).
% Innovation.
The main innovation of this work is to use concepts from analytic
combinatorics to represent reads as abstract sequences, and to specify
their generative function to estimate the probabilities of interest.
% Durability.
We provide several algorithms, which combined together give a practical
solution to the problem of calibrating seeding heuristics for short reads.
% Scope.
These results can improve current mapping algorithms and lay the
foundation of a general strategy to tackle sequence alignment problems.
\end{abstract}

@@

The rest of this document is organized to present the content in a
didactic manner. The order of the examples is meant to facilitate the
acquisition of the main concepts, meaning that we sometimes go back and
forth between different types of seeding strategies. We first give some
background on the different types of seeding methods considered here; we
then introduce the computational strategy based on transfer matrices;
finally we delve into each specific seeding method.

@@

The index is usually a hash table or a variant of the so-called
FM-index~\cite{pmid20460430}. Hash tables are typically used to index
$k$-mers, which makes them useful to search seeds of fixed size $k$
(see~\cite{pmid30346548} for a recent benchmark of $k$-mer hashing
algorithms). In contrast, some text-indexing structures have no set size
so they can be used to search seeds of different lengths. For instance,
the FM-index is a compact data structure based on the suffix
array~\cite{manber1993suffix} and on the Burrows-Wheeler
transform~\cite{burrows1994block}, emulating a suffix trie with a much
smaller memory footprint~\cite{ferragina2000opportunistic,
ferragina2005indexing}.

@@

Using the method above, we can compute the probability that a seeding
scheme using exact seeds is off target. This depends on the exact
specifications of the problem, so for illustration purposes, we choose
reads of size $k=50$ sequenced with an instrument with error rate
$p=0.01$, and we use exact seeds of size $\gamma=17$. We then vary the
number of duplicates $N$ and their divergence rate $\mu$, which both
depend on the particular read being sequenced.

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_exact.pdf}
\caption{\textbf{Off target seeding probabilities (exact seeds).}
The curves show the probability that seeding is off target for exact seeds
of size $\gamma=17$ in reads of size $k=50$ nucleotides sequenced with an
error rate $p=0.01$. Each line shows a different number of duplicate
sequences $N$ from 1 to 10 and the x-axis shows the divergence rate $\mu$,
defined as the probability that a given duplicate differs from the target
at any nucleotide.}
\label{fig:curves_exact}
\end{figure}

Fig.~\ref{fig:curves_exact} shows the result for $N$ from 1 to 10 and for
$\mu$ from 0 to 0.20. The first observation is that the probability that
seeding is off target increases with $N$. This is clearly visible from
expression (\ref{eq:Poff}). This can also be understood intuitively
because the probability of null seeding decreases when there are more
potential candidates, and since the probability of on-target seeding does
not change, the probability of off-target seeding must increase.

@@

Using the method above, we can compute the probability that a seeding
scheme using skip seeds is off target. Using the same settings as in
section~\ref{sec:illdual} ($k=50$, $p=0.01$ and $\gamma=17$), we decide to
use skip-9 seeds and vary the number of duplicates $N$ and their
divergence rate $\mu$.

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_skip.pdf}
\caption{\textbf{Off target seeding probabilities (skip seeds).}
The curves show the probability that seeding is off target for skip-9
seeds of size $\gamma=17$. Read size and error rate are the same as in
Fig.~\ref{fig:curves_exact}, \textit{i.e.} $k=50$ and $p=0.01$. Each line
shows a different number of duplicate sequences $N$ from 1 to 10 and the
x-axis shows the divergence rate $\mu$, defined as the probability that a
given duplicate differs from the target at any nucleotide.}
\label{fig:curves_skip}
\end{figure}

Fig.~\ref{fig:curves_skip} shows the result for $N$ from 1 to 10 and for
$\mu$ from 0 to 0.20. The curves have the same general aspect as those of
Fig.~\ref{fig:curves_exact}. The probability that seeding is off target
increases with $N$. There is again a  worst value of $\mu$, because the
maximum of $P(\overline{S}_0 \cap \overline{S}_1)$ minimizes expression
(\ref{eq:Poff}) for every value of $N$. But the value is different from
that of Fig.~\ref{fig:curves_exact} because $P(\overline{S}_0 \cap
\overline{S}_1)$ is computed for skip-9 seeds instead of exact seeds (here
the worst value of $\mu$ is approximately 0.07).

@@

Using the same settings as in section~\ref{sec:illdual} ($k=50$, $p=0.01$
and $\gamma=17$), we decide to use MEM seeds and vary the number of
duplicates $N$ and their divergence rate $\mu$.

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_mem.pdf}
\caption{\textbf{Off target seeding probabilities (MEM seeds).}
The curves show the probability that seeding is off target for skip-9
seeds of size $\gamma=17$. Read size and error rate are the same as in
Fig.~\ref{fig:curves_exact}, \textit{i.e.} $k=50$ and $p=0.01$. Each line
shows a different number of duplicate sequences $N$ from 1 to 10 and the
x-axis shows the divergence rate $\mu$, defined as the probability that a
given duplicate differs from the target at any nucleotide.}
\label{fig:curves_mem}
\end{figure}

Fig.~\ref{fig:curves_mem} shows the result for $N$ from 1 to 10 and for
$\mu$ from 0 to 0.20. The curves have the same general aspect as those of
Fig.~\ref{fig:curves_exact}. The probability that seeding is off target
increases with $N$, as shown by expression (\ref{eq:PMEM}).

@@

@@

We have devised a set of methods to compute the probability that seeding
heuristics fail and commit the mapping process to an error. This
fills a knowledge gap to understand and calibrate the performance of the
seeding heuristics. The pillars of our strategy are borrowed from analytic
combinatorics~\cite{flajolet2009analytic, sedgewick2013introduction}, even
though we do not follow the complete programme. Constructing generating
functions usually serves the purpose of finding their singularities in
order to approximate the solution. In our case, however, the weighted
generating functions cannot be computed so this strategy is not
applicable.

@@

Another interesting fact is that skip seeds can have better performance
than exact seeds in the sense that they can lower off-seeding
probabilities. However, this always comes at the cost of accuracy because
skipping nulceotides reduces the probability of on-target seeding.

Finally, we also observed that MEM seeds cause a very serious increase in
off-target seeding compared to exact seeds and skip seeds. This does not
mean that MEM seeding is a bad strategy (it is usually faster than the
other methods), but it is a good idea to keep the performance in check and
switch method or even skip the read altogether if the chances of
discovering the target are too low.

Regarding the methodology, the Monte Carlo approach of
algorithm~\ref{alg:mcmc} is relatively straightforward, so one may wonder
why the approach with weighted generating functions is necessary for MEM
seeds. The only reason is precision. To estimate the frequency of an event
by Monte Carlo sampling, it must occur at least a few times in the
simulation. For instance, with 10 million rounds of sampling, frequencies
around $1/1,000,000$ or lower cannot be measured accurately. When one is
interested only in frequent events, it is thus a perfectly reasonable
strategy. On the other hand, for $N < 20$, the probability that MEM
seeding is null or off-target is relatively small, so we need a method
that is accurate in this range. Fortunately, the transfer matrix method is
fast because the dimension of the matrix $\mathring{M}_N(z)$ is small and
the computations are not prohibitive.

@@

We acknowledge the financial support of the Spanish Ministry of Economy,
Industry and Competitiveness (`Centro de Excelencia Severo Ochoa
2013-2017', Plan Nacional BFU2012-37168), of the CERCA
Programme~/~Generalitat de Catalunya, and of the European Research Council
(Synergy Grant 609989). R.~C. was supported by the People Programme
(Marie Curie Actions) of the European Union's Seventh Framework Programme
(FP7/2007-2013) under REA grant agreement 608959. We also acknowledge
support of the Spanish Ministry of Economy and Competitiveness (MEIC) to
the EMBL partnership.

>>>>

\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{blkarray}
\usepackage[font=small]{caption}
\usepackage{cite}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{url}

@@

\title{Calibrating seed-based alignment heuristics with Sesame}

@@

\begin{abstract}
% Statement of the problem.
The increasing throughput of DNA sequencing technologies creates a need
for faster algorithms. The fate of most reads is to be mapped to a
reference sequence, typically a genome. Modern mappers rely on heuristics
to gain speed at a reasonable cost for accuracy. In the seeding heuristic,
short matches between the reads and the genome are used to narrow the
search to a set of candidate locations. Several seeding variants used in
modern mappers show good empirical performance but they are difficult to
calibrate or to optimize for lack of theoretical results.
% Summary of the results.
Here we develop a theory to estimate the probability that the correct
location of a read is filtered out during seeding, resulting in mapping
errors. We describe the properties of simple exact seeds, skip-seeds and
MEM seeds (Maximal Exact Match seeds).
% Innovation.
The main innovation of this work is to use concepts from analytic
combinatorics to represent reads as abstract sequences, and to specify
their generative function to estimate the probabilities of interest.
% Durability.
We provide several algorithms, which combined together give a workable
solution for the problem of calibrating seeding heuristics for short
reads. We also provide a C implementation of these algorithms in a library
called Sesame.
% Scope.
These results can improve current mapping algorithms and lay the
foundation of a general strategy to tackle sequence alignment problems.
% Availability
The Sesame library is open source and available for download at
\url{https://github.com/gui11aume/sesame}.
\end{abstract}

@@

The rest of this document is not organized by logical units but by
increasing level of difficulty. The order of the examples is meant to
ease the acquisition of the main concepts, meaning that we sometimes go
back and forth between different types of seeding strategies. We first
give some background on the different types of seeding methods considered
here; we then introduce the computational strategy based on transfer
matrices; finally we delve into each specific seeding method.

@@

The index is usually a hash table or a variant of the so-called
FM-index~\cite{ferragina2000opportunistic, ferragina2005indexing}. Hash
tables are typically used to index $k$-mers, which makes them useful to
search seeds of fixed size $k$ (see~\cite{pmid30346548} for a recent
benchmark of $k$-mer hashing algorithms). In contrast, some text-indexing
structures have no set size so they can be used to search seeds of
different lengths. For instance, the FM-index is a compact data structure
based on the suffix array~\cite{manber1993suffix} and on the
Burrows-Wheeler transform~\cite{burrows1994block}, emulating a suffix trie
with a much smaller memory footprint~\cite{ferragina2000opportunistic,
ferragina2005indexing}.

@@

We illustrate the strategy delineated above for reads of size $k=50$
sequenced with an instrument with error rate $p=0.01$, when using exact
seeds of size $\gamma=17$.

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_exact.pdf}
\caption{\textbf{Off target seeding probabilities (exact seeds).}
The curves show the probability that seeding is off target for exact seeds
of size $\gamma=17$ in reads of size $k=50$ nucleotides sequenced with an
error rate $p=0.01$. Each line shows a different number of duplicate
sequences $N$ from 1 to 10 and the x-axis shows the divergence rate $\mu$,
defined as the probability that a given duplicate differs from the target
at any given position.}
\label{fig:curves_exact}
\end{figure}

Fig.~\ref{fig:curves_exact} shows the result for a number of duplicates
$N$ from 1 to 10 and for a divergence rate $\mu$ from 0 to 0.20. The first
observation is that the probability that seeding is off target increases
with $N$. This is clearly visible from expression (\ref{eq:Poff}). This
can also be understood intuitively because the probability of null seeding
decreases when there are more potential candidates, and since the
probability of on-target seeding does not change, the probability of
off-target seeding must increase.

@@

We illustrate the strategy delineated above Using the same settings as in
section~\ref{sec:illdual} ($k=50$, $p=0.01$ and $\gamma=17$), except that
we use skip-9 seeds instead of exact seeds.

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_skip.pdf}
\caption{\textbf{Off target seeding probabilities (skip seeds).}
The curves show the probability that seeding is off target for skip-9
seeds of size $\gamma=17$. Read size and error rate are the same as in
Fig.~\ref{fig:curves_exact}, \textit{i.e.} $k=50$ and $p=0.01$. Each line
shows a different number of duplicate sequences $N$ from 1 to 10 and the
x-axis shows the divergence rate $\mu$, defined as the probability that a
given duplicate differs from the target at any given position.}
\label{fig:curves_skip}
\end{figure}

Fig.~\ref{fig:curves_skip} shows the result for a number of duplicates $N$
from 1 to 10 and for a divergence rate $\mu$ from 0 to 0.20. The curves
have the same general aspect as those of Fig.~\ref{fig:curves_exact}. The
probability that seeding is off target increases with $N$. There is again
a  worst value of $\mu$, because the maximum of $P(\overline{S}_0 \cap
\overline{S}_1)$ minimizes expression (\ref{eq:Poff}) for every value of
$N$. But the value is different from that of Fig.~\ref{fig:curves_exact}
because $P(\overline{S}_0 \cap \overline{S}_1)$ is computed for skip-9
seeds instead of exact seeds (here the worst value of $\mu$ is
approximately 0.07).

@@

We illustrate the strategy delineated above 
using the same settings as in section~\ref{sec:illdual} ($k=50$, $p=0.01$
and $\gamma=17$), except that we replace exact seed by MEM seeds.

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_mem.pdf}
\caption{\textbf{Off target seeding probabilities (MEM seeds).}
The curves show the probability that seeding is off target for MEM
seeds of minimum size $\gamma=17$. Read size and error rate are the same
as in Fig.~\ref{fig:curves_exact}, \textit{i.e.} $k=50$ and $p=0.01$. Each
line shows a different number of duplicate sequences $N$ from 1 to 10 and
the x-axis shows the divergence rate $\mu$, defined as the probability
that a given duplicate differs from the target at any given position.}
\label{fig:curves_mem}
\end{figure}

Fig.~\ref{fig:curves_mem} shows the result for a number of duplicates $N$
from 1 to 10 and for a divergence rate $\mu$ from 0 to 0.20. The curves
have the same general aspect as those of Fig.~\ref{fig:curves_exact}. The
probability that seeding is off target increases with $N$, as shown by
expression (\ref{eq:PMEM}).

@@

\section{The Sesame library}
\label{sec:sesame}

We implemented the methods and algorithms presented here in an open-source
C library to compute seeding probabilities. The library is called Sesame
and is available at \url{https://github.com/gui11aume/sesame}.

\subsection{Main features of Sesame}

Sesame contains functions to compute the seeding probabilities described
here. All the functions were tested against simulations to ensure that the
implementation is accurate and the code was checked extensively by static
analysis and unit testing.

Computating seeding probabilities can take up to a few seconds, even when
replacing iterative methods by Monte Carlo simulations. This is
incompatible with the speeed requirements of modern mappers, so Sesame has
an interface for computations that must be performed online. In this mode,
Sesame uses a type of lazy evaluation where the results are computed only
the first time, and stored in memory for reuse on subsequent calls.

Storing the results in memory is an efficient strategy because three
parameters are constant throughout the sequencing run: the minimum seed
size $\gamma$, the read size $k$ and the error rate of the sequencer $p$
(and when using skip seeds, the amount of skipping $n$ is also constant).
Only two parameters depend on the read: the number of duplicates $N$ and
their divergence rate $\mu$. In this mode, Sesame automatically switches
to Monte Carlo sampling when $N$ is large to save time. Also, the input
parameters are ``snapped'' to a predefined grid of set values for $N$ and
$\mu$, so that few computations are performed and most of the calls are
actually memory lookups. Sesame can thus be integrated in short read
mappers without being a bottleneck.

Alternatively, the probabilities of interest can be computed offline,
saved to disk and loaded at run time. This is particularly useful if the
sequencing runs follow some standard conditions with a known error rate,
because the computations can be recycled between runs.

Finally, Sesame also has an offline interface, where seeding probabilities
are computed exactly as requested by the users, \textit{i.e.} without
modifying the algorithm or the parameters, and also without storing the
results in memory.

The Sesame manual, available from the repository, contains additional
information and explains in detail how to use the library.

\subsection{Using Sesame to compare seeding strategies}

As a tool to compute seeding probabilities, Sesame can be used to compare
the merits of different strategies. The kind of insight that we can gain
from such calculations was already showcased in
Fig.~\ref{fig:curves_exact}, Fig.~\ref{fig:curves_skip} and
Fig.~\ref{fig:curves_mem}, where the numbers were computed using Sesame.

To further showcase the potential benefits of computing seeding
probabilities, we use Sesame to compare the default seeding strategies of
BWA-MEM~\cite{li2013aligning} and Bowtie2~\cite{pmid22388286}. Note that
both mappers use advanced techniques to refine the seeds, so this
comparison does not reflect the true performance of the mappers. It is
nevertheless useful to know the baseline of each strategy. The default of
BWA-MEM it to use MEM seeds of minimum size 19; that of Bowtie2 is to use
skip-9 seeds of size 16.

\begin{figure*}[t]
\begin{center}
\includegraphics[scale=.41]{mortal_kombat.pdf}
\end{center}
\caption{\textbf{Comparing seeding strategies with Sesame.} Left: MEM
seeds with $\gamma=19$ and $\mu=0.06$. Right: skip-9 seeds with
$\gamma=16$ and $\mu=0.06$. The probabilities that seeding is off target
were computed using Sesame. Each curve represents the probability for a
given number of duplicates ($N$). Estimates using iterative methods (skip
seeds and MEM sees where $N \leq 20$) were computed to within 1\%
accuracy. Estimates using Monte Carlo sampling (MEM seeds where $N > 20$)
were computed as the average of 500 million simulations.
}
\label{fig_mortal_kombat}
\end{figure*}

The probabilities that seeding is off target for different read sizes $k$
and different number of duplicates $N$ are plotted in
Fig.~\ref{fig_mortal_kombat}. The left panel shows the results for MEM
seeds and the right panel shows the results for skip seeds. Here the error
rate $p$ is set to 1\%, close to the specifications the Illumina platform,
and the divergence rate between duplicates $\mu$ is set to an arbirtarry
value of 6\%.

The behaviors of the two types of seeds are dramatically different. Let us
start with MEM seeds. For every value of $N$, the probability initially
increases with the read size, and then drops exponentially. The initial
increase is a hallmark of MEM seeds; it is due to the fact that duplicates
can mask the target. Note that the asymptotic decay depends on the number
of duplicates $N$ because each duplicate can mask the target and thereby
reduces the probability that it is discovered. Overall, these results show
that the performance of MEM seeds is poor when the target has more than
approximately 20 duplicates.

Turning to skip seeds, we see that the curves have a staircase look with a
drop every 10 nucleotides. This is so because seeding probabilities remain
unchanged until there is space for another seed of size 16 on the read.
The curves are otherwise decreasing with a steady exponential trend where
the asymptotic decay does not depend on the number of duplicates $N$. The
reason is that duplicates do not prevent the target from being discovered,
they merely fool the mapper when the target was not found. This property
makes the asymptotic decay of skip seeds substantially faster that of MEM
seeds when $N$ is large.

Those properties are self evident in retrospect, but they are not
necessarily obvious from the definitions of MEM seeds and skip seeds.
Keeping in mind that the asymptotic decay depends on $N$ for MEM seeds
helps understand what their main limitation is.

Fig.~\ref{fig_mortal_kombat} suggests that skip-9 seeds of size 16 are
just better than MEM seeds of size 19. However, the gain in sensitivity
comes at the cost of a larger candidate set, slowing down the mapping
process. To remain competitive, Bowtie2 further filters the candidate set
using a priority policy. But since some candidates are not checked, the
probability that seeding is off target is larger than shown in
Figure~\ref{fig_mortal_kombat}.

\subsection{Key insights about MEM seeds}

At least two key insights about MEM seeds can be gained from
Fig.~\ref{fig_mortal_kombat}. The first is that it is probably not worth
for a MEM-based mapper to check all the candidate loci when there are more
than approximately 20 them. The mapper may find the correct location, but
even if this is the case, the mapping quality will remain low because the
prior chances of failure were high. A better strategy is to either bail
out ot not waste time or to switch to a more sensitive seeding method (BWA
opts for the second and uses a re-seeding policy). It is also important
to note that this decision should be based on an estimated value of $N$
and not, for instance, on the size of the seed or some other variable.

A second insight is that for MEM seeds, the off-target rate is always
above $10^{-3}$ for reads of 50 nucleotides or fewer. Here it is important
to mention that the value $\mu=0.06$ is not even the worst for reads of
this size when $p=0.01$ (according to Fig.~\ref{fig:curves_mem}, the worst
value is around 0.02). So if $\mu$ is unknown and one wishes to be
conservative, it seems that reads of 50 nucleotides cannot be mapped with
confidence better than $1/1000$.

However, this is not true for the reason that it is practically impossible
to map an Illumina read of size 50 to the wrong location when $N=0$ (see
section~\ref{sec:random_seeds}). This means that tweaking the seeding
method to improve the mapping quality score will have less impact than
checking whether $N$ is 0 (in other words, whether the target is a unique
sequence).

These insights suggest that there is a way to make MEM seeds more useful
for short read mappers. We are currently working on a prototype based on
these two ideas to increase the speed and the accuracy of short read
mapping. This requires estimating $N$ and $\mu$ for every read and will be
described in a separate work.

@@

We have devised a set of methods to compute the probability that seeding
heuristics fail and commit the mapping process to an error. We have also
implemented the solution as an open source C library to perform the
computations. This fills a knowledge gap to understand and calibrate the
performance of the seeding heuristics. The pillars of our strategy are
borrowed from analytic combinatorics~\cite{flajolet2009analytic,
sedgewick2013introduction}, even though we do not follow the complete
programme. Constructing generating functions usually serves the purpose of
finding their singularities in order to approximate the solution. In our
case, however, the weighted generating functions cannot be computed so
this strategy is not applicable.

@@

Another interesting fact is that skip seeds can have better performance
than exact seeds in the sense that they can yield lower off-seeding
probabilities. However, this always comes at the cost of accuracy because
skipping nulceotides reduces the probability of on-target seeding.

Finally, we also observed that MEM seeds cause a very serious increase in
off-target seeding compared to exact seeds and skip seeds. This does not
mean that MEM seeding is a bad strategy (it is usually faster than the
other methods), but it is a good idea to keep an eye on performance and
switch method or even skip the read altogether if the chances of
discovering the target are too low.

Regarding the methodology, the Monte Carlo approach of
algorithm~\ref{alg:mcmc} is relatively straightforward, so one may wonder
why the approach with weighted generating functions is necessary for MEM
seeds. The only reason is precision. To estimate the frequency of an event
by Monte Carlo sampling, it must occur at least a few times in the
simulation. For instance, with 1 million rounds of sampling, frequencies
around $1/100,000$ or lower cannot be measured accurately. When one is
interested only in frequent events, it is thus a reasonable strategy. On
the other hand, for $N < 20$, the probability that MEM seeding is null or
off-target is relatively small, so we need a method that is accurate in
this range. Fortunately, the transfer matrix method is fast because the
dimension of the matrix $\mathring{M}_N(z)$ is small and the computations
are not prohibitive.

@@

We acknowledge the financial support of the Spanish Ministry of Economy,
Industry and Competitiveness (`Centro de Excelencia Severo Ochoa
2013-2017', Plan Estatal PGC2018-099807-B-I00), of the CERCA
Programme~/~Generalitat de Catalunya, and of the European Research Council
(Synergy Grant 609989). R.~C. was supported by the People Programme
(Marie Curie Actions) of the European Union's Seventh Framework Programme
(FP7/2007-2013) under REA grant agreement 608959. We also acknowledge
support of the Spanish Ministry of Economy and Competitiveness (MEIC) to
the EMBL partnership.
